---
title: "Tidy data wrangling"
subtitle: "demonstration"
author: "Athanasia Monika Mowinckel"
output:
  xaringan::moon_reader:
    css: [lcbc-uio.css, lcbc-uio-fonts.css, default]
    lib_dir: libs
    nature:
      titleSlideClass: [middle, right]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false 
      ratio: "16:9"
---


layout: true
    
<div class="my-sidebar"></div> 

---
name: ggplot2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      error = FALSE, 
                      warning = FALSE)

# These are the defaults
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = TRUE  #<<
)

library(knitr)
library(tidyquintro)
```
class: dark, middle, center

# ggplot2

---

.pull-left[
```{r "penguin-plot5", eval = FALSE}
ggplot(penguins,
       aes(x = bill_length_mm,
           fill = species)) +
  geom_histogram(alpha = .9, #<<
                 colour = "white")  #<<
```
]

.pull-right[
```{r "penguin-plot5-rend", echo = FALSE, ref.label="penguin-plot5"}
```
]

---

.pull-left[
```{r "penguin-plot4", eval = FALSE}
ggplot(penguins,
       aes(x = bill_length_mm,
           fill = species)) +
  geom_histogram(alpha = .8, #<<
                 colour = "white") +
  scale_fill_viridis_d() #<<
```
]

.pull-right[
```{r "penguin-plot4-rend", echo = FALSE, ref.label="penguin-plot4"}
```
]

---

.pull-left[
```{r "penguin-plot6", eval = FALSE}
ggplot(penguins,
       aes(x = bill_length_mm,
           fill = species)) +
  geom_histogram(alpha = .8,
                 colour = "white",
                 position = "dodge") + #<<
  scale_fill_viridis_d()
```
]

.pull-right[
```{r "penguin-plot6-rend", echo = FALSE, ref.label="penguin-plot6"}
```
]

---

.pull-left[
```{r "penguin-plot7", eval = FALSE}
ggplot(penguins,
       aes(x = bill_length_mm,
           fill = species)) +
  geom_density(
    alpha = .8, #<<
    colour = "white",
    position = "dodge") +
  scale_fill_viridis_d()
```
]

.pull-right[
```{r "penguin-plot7-rend", echo = FALSE, ref.label="penguin-plot7"}
```
]


---

.pull-left[
```{r "penguin-plot8", eval = FALSE}
ggplot(penguins,
       aes(y = bill_length_mm,#<<
           fill = species)) +
  geom_density(
    alpha = .8, 
    colour = "white",
    position = "dodge") +
  scale_fill_viridis_d()
```
]

.pull-right[
```{r "penguin-plot8-rend", echo = FALSE, ref.label="penguin-plot8"}
```
]

---

.pull-left[
```{r "penguin-plot9", eval = FALSE}
ggplot(penguins,
       aes(y = bill_length_mm,
           x = bill_depth_mm, 
           colour = species)) +  #<<
  geom_point() + #<<
  scale_colour_viridis_d() #<<
```
]

.pull-right[
```{r "penguin-plot9-rend", echo = FALSE, ref.label="penguin-plot9"}
```
]

---

.pull-left[
```{r "penguin-plot10", eval = FALSE}
ggplot(penguins,
       aes(y = bill_length_mm,
           x = bill_depth_mm, 
           colour = species)) +  
  geom_point() + 
  geom_line() + #<<
  scale_colour_viridis_d()
```
]

.pull-right[
```{r "penguin-plot10-rend", echo = FALSE, ref.label="penguin-plot10"}
```
]

---

.pull-left[
```{r "penguin-plot11", eval = FALSE}
ggplot(penguins,
       aes(y = bill_length_mm,
           x = bill_depth_mm, 
           colour = species)) +  
  geom_point() + 
  geom_smooth(method = "lm") + #<<
  scale_colour_viridis_d()
```
]

.pull-right[
```{r "penguin-plot11-rend", echo = FALSE, ref.label="penguin-plot11"}
```
]

---

.pull-left[
```{r "penguin-plot12", eval = FALSE}
ggplot(penguins,
       aes(y = bill_length_mm,
           x = bill_depth_mm)) +  
  geom_point(aes(colour = species)) + 
  geom_smooth(aes(colour = species),
              method = "lm") + #<<
  geom_smooth(method = "lm") + #<<
  scale_colour_viridis_d()
```
]

.pull-right[
```{r "penguin-plot12-rend", echo=FALSE, message=FALSE, warning=FALSE, ref.label="penguin-plot12"}
```
]

---
class: dark, middle, center
name: filter

# filter

---

```{r "filter1"}
filter(penguins, species == "Chinstrap")
```


---

```{r}
filter(penguins, island == "Dream")
```

---

```{r}
filter(penguins, bill_length_mm > 40)
```


---

```{r}
filter(penguins, bill_length_mm > 40, species == "Chinstrap")
```

---
class: dark, middle, center
name: select

# select

---

```{r}
select(penguins, 1, 4, 6)
```

---

```{r}
select(penguins, species, island, year, bill_length_mm)
```

---

```{r}
select(penguins, species, island, year, starts_with("bill"))
```

---

```{r}
select(penguins, species, island, body_mass_g:year, ends_with("mm"))
```

---

```{r}
select(penguins, species, ends_with("mm"))
```

---

```{r}
select(penguins, is.numeric)
```

---

```{r}
select(penguins, !is.numeric)
```

---
class: inverse, middle, center
name: pipe

# pipe

---

```{r}
penguins
```


---

```{r, eval = FALSE}
mutate(penguins, extra_col = 1)
```

same as:

```{r}
penguins |> 
  mutate(extra_col = 1)
```

---

```{r}
penguins |> 
  select(1:2, contains("bill")) |> 
  mutate(extra_col = 1) 
```

---

```{r, eval = FALSE}
penguins |> 
  select(1:2, contains("bill")) |> 
  mutate(extra_col = 1) |> 
  filter(sex == "female")
```


---
class: inverse, middle, center
name: mutate

# mutate

---

```{r}
penguins |> 
  mutate( 
    new_column = 1,
    bill_ld_ratio = bill_length_mm/bill_depth_mm
    )
```

---

```{r}
penguins |> 
  group_by(species) |> 
  mutate(
    bill_length_sp_max = max(bill_length_mm, na.rm = TRUE),
    bill_length_pc = (bill_length_mm/bill_length_sp_max)*100
    )
```


---

```{r }
penguins |> 
  mutate(across(ends_with("mm"), 
                ~ .x/10))
```


