---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyquintro)
```

# tidyquintro

<!-- badges: start -->
[![R build status](https://github.com/Athanasiamo/tidyquintro/workflows/R-CMD-check/badge.svg)](https://github.com/Athanasiamo/tidyquintro/actions)
[![Travis build status](https://travis-ci.com/Athanasiamo/tidyquintro.svg?branch=master)](https://travis-ci.com/Athanasiamo/tidyquintro)
<!-- badges: end -->

A 4 hour workshop with quick introduction to [tidyverse](https://www.tidyverse.org/).

## Workshop preparations

This package contains materials that are necessary for the workshop. 
You can install tidyquintro from [github](https://github.com/Athanasiamo/tidyquintro), which will also install the packages that the tutorial will cover, with

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("Athanasiamo/tidyquintro")
```

Loading the package will also load the necessary libraries directly
```{r, eval=FALSE}
library(tidyquintro)
```

Additionally, theere exercises connected to each of the topics which will also be accessible via the [learnr](https://rstudio.github.io/learnr/)-package.

```{r}
learnr::available_tutorials("tidyquintro")
```

The best experience with the workshop is to have an RStudio newer than 1.3. 
It is not necessary, but highly recommended.

[RStudio downloads page](https://rstudio.com/products/rstudio/download/#download)

## Timeline
All parts (except first intro) come with small exercises and breaks between.

- **Introduction to [tidyverse](https://www.tidyverse.org/) concepts** (15 minutes)  

- **Tidy data wrangling** - with translations to base-R (~ 2 hours) 
  - plotting data with [ggplot2](https://ggplot2.tidyverse.org/) (~25 min) 
  - sub-setting data with [dplyr](https://dplyr.tidyverse.org/) (~25 min)  
  - chaining commands with the pipe `%>%` (~10 min)  
  - adding and altering variables with [dplyr](https://dplyr.tidyverse.org/)  (~25 min)  

- **Lunch break** (30 min)  

- **Tidy data reshaping & summaries** - avoiding loops (~ 1.5 hours)   
  - pivoting data with [tidyr](https://tidyr.tidyverse.org/) (~25 min) 
  - grouped summaries with [dplyr](https://dplyr.tidyverse.org/) (~25 min)  
  - working with nested data using [purrr](https://purrr.tidyverse.org/) (~25 min) 
  
## Slides

All slides for the workshop can be found at:

[https://athanasiamo.github.io/tidyquintro/](https://athanasiamo.github.io/tidyquintro/)



