---
title: "Working with nested data sets"
output: 
  learnr::tutorial:
   # css: css/uio.css
    progressive: true # third level heading (###) is progressively revealed
    allow_skip: true # allows skipping exercises
runtime: shiny_prerendered
description: |
  Working on nested data is going to take time to master. This tutorial is aimed at
  showing you some of the possibilities there are when you working on such data.
---

```{r setup, include=FALSE}

library(tidyquintro)
library(learnr)
library(gradethis)

knitr::opts_chunk$set(echo = FALSE,
                 exercise.warn_invisible = FALSE)

# enable code checking
tutorial_options(exercise.checker = grade_learnr)
```


## Nesting data

Nesting data based on a grouping variable is a common way to prepare running functions on subset of the data efficiently. 
Start by grouping the data, then nesting it. Use the grouping variables species and island.

```{r nest_1, exercise = TRUE, exercise.lines = 7}
penguins %>% 
  group_by(_, _) %>% 
  nest()
```

```{r nest_1-solution}
penguins %>% 
  group_by(species, island) %>% 
  nest()
```

```{r nest_1-check}
grade_code(
  correct = random_praise(),
  incorrect = random_encouragement()
)
```

```{r nest_1-hint-1}
Did you remember to place the function first, then the colum name inside the function?
```

### Nestin data directly, without grouping first

Nesting also is possible without the explicitly group_by function first, by using `nest_by` directly.
Alter the function used below to nest directly rather than only group.


```{r nest_2, exercise = TRUE, exercise.lines = 7}
penguins %>% 
  group_by(species, island)
```

```{r nest_2-solution}
penguins %>% 
  nest_by(species, island)
```

```{r nest_2-check}
grade_code(
  correct = random_praise(),
  incorrect = random_encouragement()
)
```

```{r nest_2-hint-1}
You should use the function nest_by
```

### Alter the name of the nested column

It's not always convenient that the nested column is named "data". 
You can alter the name of the nested column using the `.key` argument.
Try calling the nested data column "measurements" in stead. 

```{r nest_3, exercise = TRUE, exercise.lines = 7}
penguins %>% 
  nest_by(species, island, .key = "_")
```

```{r nest_3-solution}
penguins %>% 
  nest_by(species, island, .key = "measurements")
```

```{r nest_3-check}
grade_code(
  correct = random_praise(),
  incorrect = random_encouragement()
)
```

```{r nest_3-hint-1}
Make sure to be careful with upper or lower case letters
```


### Adding other list columns

This style of nested data can happen because of something called "list-columns". 
As long as the output can be placed in a list, you can nest it into your data!
This is very useful when running analyses, or even making ggplots!

Try running a model predicting `body_mass_g` by sex grouped by species and island, and place the output in a column names `lm`

```{r nest_4, exercise = TRUE, exercise.lines = 7}
penguins %>% 
  nest_by(species, island) %>% 
  mutate(_ = list(lm(_ ~ _, data = _)))
```

```{r nest_4-solution}
penguins %>% 
  nest_by(species, island) %>% 
  mutate(lm = list(lm(body_mass_g ~ sex, data = data)))
```

```{r nest_4-check}
grade_code(
  correct = random_praise(),
  incorrect = random_encouragement()
)
```

```{r nest_4-hint-1}
make sure you are spelling the column names correctly
```

```{r nest_4-hint-2}
be sure that the output is placed directly within a list!
```


### Extracting summary table from model

Even better when we can grab all the summary inference statistics from the models in a neat table.
Run the same model, and use `broom::tidy` to extract the summary table from the model.
Then unnest the data based on the `tables` column.

```{r nest_g1, exercise = TRUE, exercise.lines = 7}
penguins %>% 
  nest_by(species, island) %>% 
  mutate(lm = list(lm(_ ~ _, data = _)),
         tables = list(broom::tidy(_))) %>% 
  unnest(_)
```

```{r nest_g1-solution}
penguins %>% 
  nest_by(species, island) %>% 
  mutate(lm = list(lm(body_mass_g ~ sex, data = data)),
         tables = list(broom::tidy(lm))) %>% 
  unnest(tables)
```

```{r nest_g1-check}
grade_code(
  correct = random_praise(),
  incorrect = random_encouragement()
)
```

```{r nest_g1-hint-1}
Remember to give the model output column to broom::tidy()
```

## Play around 

The best way to get a feeling for how things work is to just play around with it.
Adapt the code below and just try different things. 
See what happens, look at the possible errors etc.

```{r nest_g5, exercise = TRUE, exercise.lines = 8}
penguins %>% 
  nest_by(species, island) %>% 
  mutate(lm = list(lm(body_mass_g ~ sex, data = data)),
         tables = list(broom::tidy(lm))) %>% 
  unnest(tables)
```

<!-- ## Quiz -->

<!-- ```{r quiz} -->
<!-- quiz( -->
<!--   question("What functions can you use to subset a data set by rows?", -->
<!--     answer("dplyr's `filter()`", correct = TRUE), -->
<!--     answer("dplyr's `select()`"), -->
<!--     answer("`subset()`", correct = TRUE),
    allow_retry = TRUE -->
<!--   ), -->
<!--   question("What functions can you use to subset a data set by columns", -->
<!--     answer("dplyr's `filter()`"), -->
<!--     answer("dplyr's `select()`", correct = TRUE), -->
<!--     answer("`subset()`", correct = TRUE),
    allow_retry = TRUE -->
<!--   ), -->
<!--   question("If you want to select all columns in data 'df' that contains the string 'something', you can do that by", -->
<!--     answer("`df[grepl('something', names(df))]`", correct = TRUE), -->
<!--     answer("`select(df, starts_with('something')`"), -->
<!--     answer("`df[,'something']`"), -->
<!--     answer("`select(df, contains('something')`", correct = TRUE),
    allow_retry = TRUE -->
<!--   ), -->
<!--     question("If you want to subset rows so that you only have those below 18 years of age, you can do that by", -->
<!--     answer("`df$age < 18`"), -->
<!--     answer("`filter(df, age < 18)`", correct = TRUE), -->
<!--     answer("`df[df$age < 18,]`", correct = TRUE), -->
<!--     answer("`filter(df, age <= 18)`"),
    allow_retry = TRUE -->
<!--   )
<!-- ) -->
<!-- ``` -->
