---
title: "Workshop preparations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recommended setup
This workshop is a combination of presentation, live coding and exercises. 
If you are attending it as a webinar, having a dual monitor set-up is recommended. 
This would enable you to follow the live coding, by also typing in the commands and experience the effects you self.
Programming is a skill where you learn best by doing.

That being said, as there are also exercises you can also follow using a single monitor. 

## Install R
Make sure to have R installed on your computer.
It is recommended that you have a newer version of R installed, with minimum version of 4.1.0.
Download R from [the R-project](https://www.r-project.org/)

#### Windows computers
The tidyverse packages require [rtools](https://cran.r-project.org/bin/windows/Rtools/) to be installed. 
Make sure to have the required version installed before proceeding.

## Install RStudio

The best experience with the workshop is to have an RStudio newer than 1.3. 
It is not necessary, but highly recommended.

[RStudio downloads page](https://rstudio.com/products/rstudio/download/#download)


## Install course materials

This package contains materials that are necessary for the workshop. 

You can install tidyquintro from [my r-universe](https://drmowinckels.r-universe.dev), which will also install the packages that the tutorial will cover, with

```r
# Enable repository from drmowinckels
options(repos = c(
  drmowinckels = 'https://drmowinckels.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
  
# Download and install tidyquintro in R
install.packages('tidyquintro')
```

### Loading the package
Loading the package will also load the necessary libraries directly
```r
library(tidyquintro)
```

Additionally, there are exercises connected to each of the topics which will also be accessible via the [learnr](https://rstudio.github.io/learnr/)-package.

```{r}
learnr::available_tutorials("tidyquintro")
```

#### If install fails
If the above install fails, you can try installing everything individually and then the package from source.
Start by installing the dependent packages for the course:

```r
install.packages(c("tidyverse", "palmerpenguins", "learnr", "remotes"))
remotes::install_github("rstudio-education/gradethis")
``` 

When these are successful, you should be able to install the course package.

If you are still struggling to install the package, please create [a ticket](https://github.com/Athanasiamo/tidyquintro/issues) pasting the command and output you get so I can help you get it sorted.
